\textbf{\Large Computer Vision-based System for Impaired Human Vision Compensation}

### 2019-01-15 bare_conf.tex
more fixes according to review.

\begin{abstract}
%\boldmath
This paper presents a high-level architecture for computer vision-based system for partial compensation of lost or impaired human vision. It combines standard smartphone device, external deep learning-based image processing infrastructure and audio/tactile user interface. The proposed architecture is based on  input from user-centered design process, involving end users into system development. The paper discusses user needs and expectations towards electronic travelling aids for the blind and highlights limitations of the existing solutions. The suggested architecture may be used as a basis for developing computer vision-based tools for visually impaired individuals. 
%In this short article we describe and discuss high level architecture of computer vision-based system for partial compensation of lost or impaired human vision. Combining standard smartphone device, external deep learning-based live video processoing infrastructure and audio/tactile user interface into compact visual perception compensation device .
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

\begin{IEEEkeywords}
computer vision; deep learning; mobile application; aid for blind and visually impaired; audio feedback; tactile feedback; impaired human vision compensation; user-centered design.%
\end{IEEEkeywords}
% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle
\section{Introduction}
\label{sec:introduction}
More than $250$ million people worldwide have moderate to severe vision impairment, while ($\approx 36$ million are blind) \cite{Bourne}. During the past decades significant effort was devoted to develop computer vision and other sensor based aids (e.g. \cite{Caraiman}, \cite{Csapo}, \cite{Poggi}, \cite{Zientara}) for helping the blind and visually impaired users to perceive the surrounding world better. However, computer vision is a rapidly evolving field; systems based on the aforementioned approaches become outdated in a relatively short time and often lack accuracy and reliability in real-world conditions in comparison to the state-of-the-art technologies. In this article, we describe a system, which employs modern computer vision techniques for compensation of lost or impaired vision function in humans. Many of the previously proposed electronic aids for the blind count on highly specialized hardware, for instance smart glasses \cite{Zientara}, Microsoft Kinnect sensors \cite{Owayjan}, helmet-mounted photo sensors and cameras \cite{Dunai}. Such systems often lack convenience and bring additional complexity into daily tasks, which visually impaired people currently manage by the help of a white cane. This project puts major emphasis on developing a tool, which seamlessly integrates with the hardware visually impaired persons already have on-hand and are used to (smartphone). This paper describes a high level architecture of the technology aid for visually impaired people utilizing the latest achievements in the field of computer vision.

% corr start
This paper is structured as follows. We start our presentation with Section~\ref{sec:method}, which describes the methodological foundations of this research. Section~\ref{sec:results} detalizes the architecture of the aforementioned system. Finally, Section~\ref{sec:discussion} summarizes and discusses the results of conducted analysis.
% corr end

@@

Following the principles of the participatory design \cite{Kensing}, \cite{Carroll} we included the end-users into the design process of the system. Two visually impaired persons participated in discussion and focus group meetings together with the project team to identify the key properties of the solution. During the six discussion sessions major attention was paid to:
\begin{enumerate}
\item Functionality and key features of the solution addressing real-life scenarios of visually impaired persons;
\item Technical feasibility, connecting user needs to the latest development in computer vision and assistive technologies;
\item Interface between the visually impaired person and the assistive technology;
\item Appearance, usability and potential costs of the proposed product. 
\end{enumerate}
Participatory design process included testing of assistive technologies for the blind and visually impaired people using widely available mobile phone apps like Seeing AI, BlindWays, Be My Eyes, Aipoly Vision, TapTapSee, etc.
%Design process included testing of the existing mobile phone apps (Seeing AI, BlindWays, Be My Eyes, Aipoly Vision, TapTapSee) for assisting visually impaired people. 
Semi-structured tests were performed by visually impaired participants and researchers from the project group. Strengths and weaknesses of the existing solutions are reflected in the design of the proposed system.

@@

The key part of the proposed system is a smartphone device, connecting user interface (UI) and data processing components into a well-functioning ecosystem. Smartphone was selected due to its high availability and capability to perform required computations or forward visual information for processing to the external server via the Internet.

@@

% User interface
UI consists of a wearable forehead belt, integrating RGB and depth cameras, inertial measurement unit (IMU), tactile feedback device and bone conduction headphones. Although more expensive than standard ones, bone conduction headphones do not block the ears and allow the user to perceive audio information from the surrounding environment and the system at the same time, which is essential for the visually impaired people.

@@

%The tactile feedback may improve the quality of users perception (e.g. by ). UI can interface smartphone component via USB or Wifi connection.

% External vision processor
%visus sutrumpinimus reikia pirma parasyt pilnai, pvz Neural Network (NN)
Although modern high-end smartphones are powerful computational devices, many computer vision algorithms  require hardware of significantly higher processing power (e.g. a server equipped with a high-end graphics processing unit (GPU)). Therefore, an external server is used for live vision data processing. In our design we envision a computational server hosting a set of computer vision algorithms implemented and exposed via web services: faster regions with convolutional neural network features (faster R-CNN) object detector \cite{Ren}, trained to detect important objects, convolutional and recurrent neural network-based (CNN and RNN) scene descriptor \cite{Liu}, which provides textual annotation of a given RGB image, place recognition \cite{Ohn-Bar}, face detection and recognition \cite{Amos}, obstacle detection \cite{Laina}, action recognition \cite{Valor}, among others. After video analysis is completed on the server side, results are transmitted back to the smartphone component and represented to the user through the aforementioned UI.

% How user controls system
The user interacts with the system through gestures, voice commands or specifically designed user interface on the smartphone. User input is required for selecting operation mode of the system, i.e. switching between optical character recognition (OCR), currency recognition, object detection, navigation aid and other available functions.

@@

Discussions on the key features of the proposed solution revealed a gap between the functionality of the existing tools, which are often focused on advanced navigation and scene description features, and actual user needs. For example, the end-users highlighted that a simplistic solution for navigating is lacking. Major focus was put on an easy-to-use and minimalistic tool, which can be trusted. Reliable object detection, direction and approximate distance to an obstacle were highlighted as the main requirements. To our surprise, only 5-8 distinctive object classes (for instance, buss stop, pedestrian crossing, doors, stairs, etc) were of major importance while navigating. Additional information, such as type of a passing car, blossoming flowers in the park or a person riding a bicycle was considered as overwhelming and distractive.

%Povilai, paziurek ar cia viskas OK, pridek ka nors nuo saves
% Pamineti mobilenet ir pan.
Computer vision technology is under rapid development and during the last years made a major breakthrough in performance and efficiency. Deep learning neural networks, which are de facto standard in modern image/video processing in many real-world problems allow to achieve accuracies, similar to that of human decision (e.g. face recognition \cite{Amos}, object detection \cite{Ren}, among others). These models are now well supported by software libraries (e.g. \cite{Tensorflow}) and dedicated processing hardware is built-in even in relatively low-power computational devices, such as smartphones. These devices are also equipped with 4G Internet connectivity and sufficient computational power to perform partial or in certain cases even full vision data processing locally. The maturity of the aforementioned technologies suggest that our proposed computational aid for visually impaired persons may be highly feasible.

@@

Both open source (e.g. \cite{Tensorflow}) and state of the art commercial computer vision software libraries (e.g. \cite{Verilook}) may be applied implementing the suggested architecture. Moreover, detalization and specification of the suggested system's hardware components following open hardware approach - by providing open repository with the detailed list of electronics components, schematics, and 3D models of mechanical parts, could provide a solid basis for semi-standard reference platform for the researchers in the field.

It is important to emphasize that the suggested system does not aim to replace the main travelling aid of visually impaired people - the white cane. Instead, this computer vision tool aims to enhance and push the perception of the surrounding environment boundary from 1-2 meters (achieved by using the white cane) to 2-10 meters. It may improve route planning and identification of objects of interest (for instance, doors, stairs, elevators, bus stops, etc.). This functionality corresponds with the main requirement for electronic travelling aids highlighted by the end users - direction and distance estimation to a selected object. The proposed hardware architecture can be used as basis for various computer vision-based software modules, aiming to assist visually impaired users in daily activities (e.g. outdoor/indoor navigation, object/face recognition, obstacle detection, etc.).

>>>>

\begin{abstract}
%\boldmath
This paper presents a high-level architecture for computer vision-based system for partial compensation of lost or impaired human vision. It combines standard smartphone device, external deep learning-based image processing infrastructure and audio/tactile user interface. The proposed architecture is based on  input from user-centered design process, involving end users into system development. The paper discusses user needs and expectations towards electronic travelling aids for the blind and highlights limitations of the existing solutions. The suggested architecture may be used as a basis for developing computer vision-based tools for visually impaired individuals. 
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

\begin{IEEEkeywords}
computer vision; deep learning; mobile application; aid for blind and visually impaired; audio feedback; tactile feedback; impaired human vision compensation; user-centered design.%
\end{IEEEkeywords}
% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle
\section{Introduction}
\label{sec:introduction}
More than $250$ million people worldwide have moderate to severe vision impairment, while ($\approx 36$ million are blind) \cite{Bourne}. During the past decades significant effort was devoted to develop computer vision and other sensor based aids (e.g., \cite{Caraiman}, \cite{Csapo}, \cite{Poggi}, \cite{Zientara}) for helping the blind and visually impaired users to perceive the surrounding world better. However, computer vision is a rapidly evolving field; systems based on the aforementioned approaches become outdated in a relatively short time and often lack accuracy and reliability in real-world conditions in comparison to the state-of-the-art technologies. In this article, we describe a system, which employs modern computer vision techniques for compensation of lost or impaired vision function in humans. Many of the previously proposed electronic aids for the blind count on highly specialized hardware, for instance smart glasses \cite{Zientara}, Microsoft Kinnect sensors \cite{Owayjan}, helmet-mounted photo sensors and cameras \cite{Dunai}. Such systems often lack convenience and bring additional complexity into daily tasks, which visually impaired people currently manage by the help of a white cane. This project puts major emphasis on developing a tool, which seamlessly integrates with the hardware visually impaired persons already have on-hand and are used to (smartphone). This paper describes a high level architecture of the technology aid for visually impaired people utilizing the latest achievements in the field of computer vision.

% corr start
This paper is structured as follows. We start our presentation with Section~\ref{sec:method}, which describes the methodological foundations of this research. Section~\ref{sec:results} focuses on the architecture of the suggested vision compensation system. Finally, Section~\ref{sec:discussion} summarizes and discusses the results of the conducted analysis.
% corr end

@@

Following the principles of the participatory design \cite{Kensing}, \cite{Carroll} we included the end-users into the design process of the system. Two visually impaired persons participated in discussion and focus group meetings together with the project team to identify the key properties of the solution. During the six discussion sessions, major attention was paid to:
\begin{enumerate}
\item Functionality and key features of the solution addressing real-life scenarios of visually impaired persons;
\item Technical feasibility, connecting user needs to the latest development in computer vision and assistive technologies;
\item Interface between the visually impaired person and the assistive technology;
\item Appearance, usability and potential costs of the proposed product. 
\end{enumerate}
Participatory design process included testing of assistive technologies for the blind and visually impaired people using widely available mobile phone apps like Seeing AI, BlindWays, Be My Eyes, Aipoly Vision, TapTapSee, etc.
%Design process included testing of the existing mobile phone apps (Seeing AI, BlindWays, Be My Eyes, Aipoly Vision, TapTapSee) for assisting visually impaired people. 
Semi-structured tests were performed by visually impaired participants and researchers from the project group. Strengths and weaknesses of the existing solutions are reflected in the design of the proposed system.

@@

The key part of the proposed system is a smartphone device, connecting User interface (UI) and data processing components into a well-functioning ecosystem. Smartphone was selected due to its high availability and capability to perform required computations or forward visual information for processing to the external server via the Internet.

@@

% User interface
UI consists of a wearable forehead belt, integrating RGB and depth cameras, Inertial measurement unit (IMU), tactile feedback device and bone conduction headphones. Although more expensive than standard ones, bone conduction headphones do not block the ears and allow the user to perceive audio information from the surrounding environment and the system at the same time, which is essential for the visually impaired people.

@@

% External vision processor
%visus sutrumpinimus reikia pirma parasyt pilnai, pvz Neural Network (NN)
Although modern high-end smartphones are powerful computational devices, many computer vision algorithms  require hardware of significantly higher processing power (e.g., a server equipped with a high-end Graphics processing unit (GPU)). Therefore, an external server is used for live vision data processing. In our design we envision a computational server hosting a set of computer vision algorithms implemented and exposed via web services: faster Regions with convolutional neural network features (faster R-CNN) object detector \cite{Ren}, trained to detect important objects, convolutional and recurrent neural network-based (CNN and RNN) scene descriptor \cite{Liu}, which provides textual annotation of a given RGB image, place recognition \cite{Ohn-Bar}, face detection and recognition \cite{Amos}, obstacle detection \cite{Laina}, action recognition \cite{Valor}, among others. After video analysis is completed on the server side, results are transmitted back to the smartphone component and represented to the user through the aforementioned UI.

% How user controls system
The user interacts with the system through gestures, voice commands or specifically designed user interface on the smartphone. User input is required for selecting operation mode of the system, i.e., switching between Optical character recognition (OCR), currency recognition, object detection, navigation aid and other available functions.

@@

Discussions on the key features of the proposed solution revealed a gap between the functionality of the existing tools, which are often focused on advanced navigation and scene description features, and actual user needs. For example, the end-users highlighted that a simplistic solution for navigating is lacking. Major focus was put on an easy-to-use and minimalistic tool, which can be trusted. Reliable object detection, direction and approximate distance to an obstacle were highlighted as the main requirements. To our surprise, only 5-8 distinctive object classes (for instance, buss stop, pedestrian crossing, doors, stairs, etc.) were of major importance while navigating. Additional information, such as type of a passing car, blossoming flowers in the park or a person riding a bicycle was considered as overwhelming and distractive.

%Povilai, paziurek ar cia viskas OK, pridek ka nors nuo saves
% Pamineti mobilenet ir pan.
Computer vision technology is under rapid development and during the last years made a major breakthrough in performance and efficiency. Deep learning neural networks, which are de facto standard in modern image/video processing in many real-world problems allow to achieve accuracies, similar to that of human decision (e.g., face recognition \cite{Amos}, object detection \cite{Ren}, among others). These models are now well supported by software libraries (e.g., \cite{Tensorflow}) and dedicated processing hardware is built-in even in relatively low-power computational devices, such as smartphones. These devices are also equipped with 4G Internet connectivity and sufficient computational power to perform partial or in certain cases even full vision data processing locally. The maturity of the aforementioned technologies suggest that our proposed computational aid for visually impaired persons may be highly feasible.

@@

Both open source (e.g., \cite{Tensorflow}) and state of the art commercial computer vision software libraries (e.g., \cite{Verilook}) may be applied implementing the suggested architecture. Moreover, detalization and specification of the suggested system's hardware components following open hardware approach - by providing open repository with the detailed list of electronics components, schematics, and 3D models of mechanical parts, could provide a solid basis for semi-standard reference platform for the researchers in the field.

It is important to emphasize that the suggested system does not aim to replace the main travelling aid of visually impaired people - the white cane. Instead, this computer vision tool aims to enhance and push the perception of the surrounding environment boundary from 1-2 meters (achieved by using the white cane) to 2-10 meters. It may improve route planning and identification of objects of interest (for instance, doors, stairs, elevators, bus stops, etc.). This functionality corresponds with the main requirement for electronic travelling aids highlighted by the end users - direction and distance estimation to a selected object. The proposed hardware architecture can be used as basis for various computer vision-based software modules, aiming to assist visually impaired users in daily activities (e.g., outdoor/indoor navigation, object/face recognition, obstacle detection, etc.).
